An Invariant Framework for Conducting Reproducible Computational Science

### 2015-03-30 evolution.tex
Modification (according to dthain's suggestions)
proof reading

Given the many challenges of reproducing HEP applications, we now describe an invariant framework, which if present within large collaborations, can enable application developers to share their 
application with other researchers, and for other researchers to reproduce the shared application. 
To satisfy invariance, the framework must include mechanisms for:

@@

\item {\bf Capturing Dependencies and Configurations:} Capturing tools must record dependencies that are used by the program, including hardware, OS, kernels, static and dynamic dependencies, local and networked dependencies, source codes and data files. Stateful interactions with commercial software, such as proprietary databases and which cannot be captured due to licensing agreements must be persisted so that they can replayed later without the presence of the commercial software.  In effect, a captured application should behave exactly the same way as the application developer intended it to be. 
% In audit phase, PTU captures the execution of the application into a package. The resulting package contains the source code, its dependencies (system files, and files on network such as from CernVM-FS) and the data consumed and written by the application.

\item{\bf Preservation of Captured Entities:} By preservation we imply appropriate mechanisms for (a) documentation of the application development, and (b) automation of any task that becomes inevitably 
necessary for repeating the application in exactly the same way as the developer executed it.

Documentation and specification during application development can be an onerous one. The preservation framework must make programming tools available that 
focus less on documentation, but script, integrate and execute the dependencies so that they are resolved as part of documentation. 
Automation can extend to various tasks that are necessary for ensuring repeatability such as building software, provisioning of hardware, 
validation of software against security fixes, new features, and even monitoring the reproducibility state of a preserved application, i.e., its source code, dependencies, environment, and platform.
Automated builds and provisioning and continuous integration service can significantly lower the barrier to run the application in a new environment.

In spite of the preservation mechanisms, the application software may not run as intended. For a reproducer's understanding, it may also be useful to include 
a \emph{logical preservation unit} (PLU) that consists of a minimal execution of the software using a small, test data sample, and with specified outputs.
The provenance of this PLU must be captured so that the reproducer can make comparison with future reproduction-validation runs.

@@

%\section{Evolving the Artifact}
%
%\begin{figure*}[t]
%\centering
%\includegraphics[width=.8\textwidth]{version-evolution.eps}
%\caption{Version Evolution}
%\label{fig:version-evolution}
%\end{figure*}
%
%It is clear that the artifact, as provided, is not in a suitable form
%for preservation.  While it might be technically possible to automatically
%capture the entire virtual machine and all of the connected filesystems,
%it would require 166.8TB of storage, which would be prohibitively expensive
%for capturing this one application alone.  Further, if multiple
%similar applications are preserved, we would miss the opportunity to identify
%common dependencies and store them once for multiple artifacts.
%A more structured approach to dependency management is needed.
%
%Figure~\ref{fig:version-evolution} shows how we have evolved this artifact
%through several stages which make it more suitable for preservation.
%In each step of evolution, we make the dependencies of the artifact
%more explicit and available for analysis and automated processing.
%As noted in the previous section, the original author provided us with
%prose instructions by email which we translated into an
%executable script.  The executable script has embedded in it
%a number of external identifiers such as URLs pointing to repositories
%and paths to networked filesystems.  As a general programming practice,
%embedding such constants into the middle of a program is unwise, and so
%we extract all of those identifiers and place them \emph{outside} the
%script in a \emph{dependency map} or just \emph{map} for short.
%The dependency map lists all of the external dependencies of the application, indicating
%the type, how they are accessed, and where they are currently located.
%The resulting script then simply refers to abstract file locations such
%as \verb$GIT$ and \verb$CVMFS$, while the map file indicates where they
%are currently located. If properly constructed, the script should not refer to any external
%resource unless it is indicated in the dependency map.  We call this idealized
%artifact an \emph{abstract script}.
%
%By extracting the dependencies into the dependency map,
%we introduce great freedom for the curator to move, transform, and otherwise
%manipulate the dependencies of the artifact without damaging the artifact itself.
%A Figure~\ref{fig:version-evolution} shows, it is straightforward for an
%automated tool to examine
%all of the dependencies in the map, download those that are missing,
%and then modify the map to point to the local copies of the dependencies.
%If we group the script, dependency map, and dependencies into a \emph{package},
%we now have a self-contained artifact that can be moved from place to place.
%In some cases, it may be safe to allow the dependency map to refer to 
%trusted remote repositories.  Whether this is advisable is a judgment that
%must be made by the user or the curator, taking into account the long-term
%stability of said repositories.
%
%However, only having the package and the dependency map is not enough. The successful execution of the application on the original machine also relies on the configuration of environment variables.
%We collected the environment variables of the original machine and transformed it into one executable script. If another researcher wants to repeat the application, this executable script will be first executed. 
%The environment variables refer to the dynamic named values defined in the range of one computer, like \verb!$PATH! and \verb!$PYTHONPATH!.
%However, the dependency map refers to the actual storage location of path variables used in one abstract script. For example, the actual storage location of the path variable \verb!CVMFS! 
%is {\tt /data/cvmfs/grid.cern.ch}.
%
%\begin{figure}
%\centering
%\includegraphics[width=.5\textwidth]{solution3.eps}
%\caption{Relationship of Roles}
%\label{fig:solution3}
%\end{figure}
%
%The relationship of different roles involved in the application preservation and
%reproduction is shown in Figure~\ref{fig:solution3}.  The original author uses the packaging utility to generate the package for one application.
%Then the package, together with
%its map file and description file will be published. When
%another researcher wants to repeat the application, one copy of the package will be downloaded into the new machine and the application can be repeated.
%
%When we try to repeat one application on one new machine, one map file is necessary for the relocation of the data access targets, as
%show in Figure~\ref{fig:version-evolution}. 
%The map file clearly defines the real location of each dependency in the format of dependency variables - the real location of dependency variable {\tt GIT} is \path{/data/git/cms-ttH} and the real location of dependency variable {\tt CVMFS}
%is {\tt /data/cvmfs/grid.cern.ch}.
%The script only refers to the dependency variables defined in the map file.
%This design decouples the application script and the actual data access targets, which minimizes the impact of the evolution of different data dependencies
%and ensures the transparent access.
%The modification of the package only introduces the minimal changes of the map file on the client side.
%
%This basic approach to dependency management is a step in the right direction
%for dependencies that are \emph{explicit} and \emph{external} to the
%user's native execution environment.  However, it leaves two other problems unsolved.
%
%First, the basic approach requires that someone be \emph{aware} of the dependencies,
%whether it be the end user, the system administrator, or the archive curator.
%It seems reasonable to expect the user to be aware of a large dependency mentioned in a top-level script.
%But, oftentimes the dependency is embedded invisibly deep within the software stack,
%or is connected to the machine by the local system administrators.  No single party
%is likely to have complete information about all of the dependencies.
%Second, the basic approach assumes that the entire dependency is actually
%consumed by the artifact.  As we have suggested above (and will show below),
%this sort of program often only consumes a small fraction of what it does declare
%as a dependency.
%
%To address both of these problems, users and curators alike need tools that
%will automatically observe the dependencies of complex applications, to
%facilitate automatic and efficient preservation.
%

>>>>

Given the many challenges of reproducing HEP applications, we now describe an invariant framework. If present within large collaborations, this framework can enable application developers to share their 
application with other researchers, and for other researchers to reproduce the shared application. 
To satisfy invariance, the framework must include mechanisms for:

@@

\item {\bf Capturing Dependencies and Configurations:} 
Capturing tools must record dependencies that are used by the program, including hardware, OS, kernels, static and dynamic dependencies, local and networked dependencies, source codes and data files. Stateful interactions with commercial software, such as proprietary databases and which cannot be captured due to licensing agreements must persist such that replaying later may be accomplished without the presence of the commercial software. In effect, a captured application should behave in exactly the way the application developer intended.
% In audit phase, PTU captures the execution of the application into a package. The resulting package contains the source code, its dependencies (system files, and files on network such as from CernVM-FS) and the data consumed and written by the application.

\item{\bf Preservation of Captured Entities:} By preservation we define appropriate mechanisms for (a) documentation of the application development, and (b) automation of any task that becomes necessary to the repetition of the application in exactly the the way the application developer intended.

Documentation and specification during application development can be onerous. The preservation framework must make programming tools available that focus less on documentation, and more on scripts, integration, and execution of the dependencies such that they are resolved as part of documentation. Automation can extend to various tasks necessary for ensuring repeatability such as building software, provisioning of hardware, validation of software against security fixes, new features, and even monitoring the reproducibility state of a preserved application, i.e., its source code, dependencies, environment, and platform. Automated builds and provisioning and continuous integration service can significantly lower the barriers to running applications in a new environment.

Despite preservation mechanisms, the application software may not run as intended. For a reproducer’s understanding, it may also be useful to include a \emph{logical preservation unit} (PLU) that consists of a minimal execution of the software using a small, test data sample and with specified outputs. The provenance of this PLU must be captured so that the reproducer can compare the current run with future reproduction-validation runs.

@@

### 2015-03-30 measure.tex
proof reading

Parrot is a virtual filesystem access tool which
is used to attach applications to a variety of remote I/O systems such as HTTP, FTP, and CVMFS. It works by trapping system calls through the {\tt ptrace} interface,
replacing selected operations with remote accesses.  As a side-effect, Parrot is
also able to modify the filesystem namespace in arbitrary ways according to user
needs.  Parrot is particularly used in the high energy physics community
to provide remote access to application software via CVMFS.
To capture dependencies, we made small modifications to Parrot to record
the logical name of every file accessed by an application into an external
dependency list.  After execution is complete, a second tool is used to copy
all of the named dependencies into a package.

PTU~\cite{pham2014framework} is designed to create a package of an application by recording all of the binaries, libraries, scripts, data files, and environment variables used by a program.   PTU uses the CDE technology to observe system calls, but takes a snapshot of every file at the point of access.  In addition to files, PTU records \emph{metadata} about the execution environment, such as kernel versions, application versions, and dynamic library versions by using standard Unix commands.  PTU also records \emph{provenance} in the form of a graph that describes how each file is created or consumed by processes within the application.  Because PTU is focused solely on the problem of preservation, it can achieve lower overhead than Parrot when remote data access is not a requirement, as we show below.

@@

There are many possible heuristics for creating the preserved package
from the dependency list.  We have implemented three:
In a {\bf shallow copy}, we copy only
the exact dependencies.  Where a directory was listed, a directory is created and populated with empty files as placeholders to facilitate a directory listing.  In a {\bf medium copy},every file in every directory mentioned by the application is preserved one level deep.  In a {\bf deep copy}, every file in every directory mentioned by the application is copied recursively.  Obviously, the more aggressive the preservation, the larger the package, but the greater possibility that it can be adapted to other uses.  Other approaches to package generation might including generating the union
of multiple dependency sets, or allowing the expert user to manually add
and remove dependencies.

@@

{\bf Distribution and Deployment.}
Once generated, application packages must be collected, curated, and made
available through publically shared repositories.  There currently exist a
wide variety of services and efforts to share binary objects in this way,
and so we assume such multiple services will be available in the future.

Of more immediate interest is the ability to deploy a preserved
package at a desired execution site.  Again, long-term preservation
requires artifacts that are independent of any particular technology,
so the package must be sufficiently self-describing
to work with multiple technologies.  The packages produced by both Parrot and PTU
can today be re-executed through any of the following mechanisms:

@@

 % Docker images
\vspace{5pt}
\noindent{\bf Linux Containers and Docker Images:} Linux Containers provide multiple isolated execution instances on top of the same kernel through OS-level virtualization. Thus they can be used to persist the captured packages into images. 
Using such containers, Docker now allows to preserve the image in a more user-friendly way. 
Docker also provides portable deployment of containers across platforms, documentation of packages in a scriptable format, and versioning of container images. The image can be preserved along with dockerfiles, which are text files that contain all the commands to build a Docker image: the computational part is similar to shell scripts and can help provision similar to other provisioning tools (e.g. Chef, Puppet) and the text part is for human consumption and more suited for use with a version management system such as subversion or git, which can track any changes made to the Dockerfile. Thus dockerfiles can be used to preserve the namelist of Parrot packages and provenance description in PTU. 
Docker is integrated with a continuous build environment which will check and validate the version of the software being used, and use a more recent version to build application software.

>>>>

Parrot is a virtual filesystem access tool which
is used to attach applications to a variety of remote I/O systems such as HTTP, FTP, and CVMFS. It works by trapping system calls through the {\tt ptrace} interface,
replacing selected operations with remote accesses.  As a side-effect, Parrot is
also able to modify the filesystem namespace in arbitrary ways according to user
needs.  Parrot is particularly used in the high-energy physics community
to provide remote access to application software via CVMFS.
To capture dependencies, we made small modifications to Parrot to record
the logical name of every file accessed by an application into an external
dependency list.  After execution is complete, a second tool is used to copy
all of the named dependencies into a package.

PTU~\cite{pham2014framework} is designed to create a package of an application by recording all of the binaries, libraries, scripts, data files, and environment variables used by a program.   PTU uses the CDE technology to observe system calls, but takes a snapshot of every file at the point of access.  In addition to files, PTU records \emph{metadata} about the execution environment, such as kernel versions, application versions, and dynamic library versions by using standard Unix commands.  PTU also records \emph{provenance} in the form of a graph that describes how each file is created or consumed by processes within the application.  Because PTU is focused solely on the problem of preservation, therefore it can achieve lower overhead than Parrot when remote data access is not a requirement, as we show below.

@@

Many possible heuristics exist for creating the preserved package
from the dependency list.  We have implemented three:
In a {\bf shallow copy}, we copy only
the exact dependencies.  Where a directory was listed, a directory is created and populated with empty files as placeholders to facilitate a directory listing.  In a {\bf medium copy},every file in every directory mentioned by the application is preserved one level deep.  In a {\bf deep copy}, every file in every directory mentioned by the application is copied recursively.  Obviously, the more aggressive the preservation, the larger the package, but also the greater the possibility that the package can be adapted to other uses.  Other approaches to package generation might including generating the union
of multiple dependency sets, or allowing an expert user to manually add
and remove dependencies.

@@

{\bf Distribution and Deployment.}
Once generated, application packages must be collected, curated, and made
available through publically shared repositories.  Currently, a 
wide variety of services and efforts exist in order to share binary objects in this way,
and so we assume such multiple services will be available in the future.

Of more immediate interest is the ability to deploy a preserved
package at a desired execution site.  Again, long-term preservation
requires artifacts that are independent of any particular technology,
so the package must be sufficiently self-describing in order
to work with multiple technologies.  The packages produced by both Parrot and PTU
can today be re-executed through any of the following mechanisms:

@@

 % Docker images
\vspace{5pt}
\noindent{\bf Linux Containers and Docker Images:} 
Linux Containers provide multiple isolated instances of execution on top of the same kernel through OS-level virtualization. Thus they can be used to persist the captured packages into images. Using such containers, Docker now allows for the preservation of the image in a more user-friendly way. Docker also provides portable deployment of containers across platforms, documentation of packages in a scriptable format, and versioning of container images. The image can be preserved along with dockerfiles, which are text files that contain all the commands to build a Docker image. Similar to shell scripts, the computational section can help other provisioning tools (e.g. Chef, Puppet).  The text section, written for for human consumption, is more suited for use with a version management system such as subversion or git, which can track any changes made to the Dockerfile. Thus dockerfiles can be used to preserve the namelist of Parrot packages and provenance description in PTU. Docker is integrated with a continuous build environment that will not only check and validate the version of the software in present use, but also make use of a more recent version to build application software.
